
---
title: Embedded HTML file
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: false
---

### Using Table of Contents
Another way of organizing the page is using a table of contents which appears in a side-bar on the left-hand margin. Notice that there is a line automatically created for each heading.
The top of this embedded html file contains the following yaml at the beginning:

#### YAML code


```{r, eval = FALSE}
---
title: Embedded HTML file
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: false
---

```

#### Note of Caution
Using a table of contents does not work really well with tabs and some other elements. (That is why it was demonstrated in this embedded html file 
<span style="font-size: 20px;">&#x1F609;</span>)

### Other stuff
This section is here just to have more content. Notice how the webpage scrolls within the window while the table of contents remains fixed.

```{r}
library(data.table)
library(htmlTable)
htmlTable(data.table(iris)[, list(Samples = .N, Avg_Length = mean(Sepal.Length), Avg_Width = mean(Sepal.Width)), keyby = list(Species)],
 rnames = F, align = "lr", css.cell = "padding-left: 3em; padding-right: 3em;")
plot(iris$Sepal.Length, iris$Sepal.Width)

```

