---
title: "Maps"
output: html_document
---

## Testing mapping functions

```{r, echo=FALSE,message=FALSE,warning=FALSE}
# Plotting methods
library(data.table)
library(googleVis)
library(plotGoogleMaps)
library(ggmap)
library(ggplot2)
library(leafletR)
library(RgoogleMaps)


source("/work/cmcneil/projects/miscellaneous/Personal/General Code/MarkdownFunctions.R")


# Standard data to test (try to turn Category into color and Speed_kt into size)
data(Andrew)  # from googleVis
dt <- data.table(Andrew)
dt$Color <- c("red","blue","green")[as.numeric(dt$Category)]
setnames(dt,"Date/Time UTC","DateTime")
dt$Index <- 1:nrow(dt)
```

Different options:
`r html_list(c("gvisMap from googleVis",
               "leaflet from leafletR",
               "plotGoogleMaps from plotGoogleMaps",
               "ggmap from ggmap"))`
               
Desired functionality:
`r html_list(c("Interactive",
               "Aesthetics",
               "Markdown compatible",
               "Riley2 compatible",
               "Dynamic"),
             details=list(c("Zoom, scroll, explore area"),
                          c("size, color, shape","points, lines, areas"),
                          NULL,
                          NULL,
                          c("Like D3 or Shiny")),kind="dl")`

### gvisMap from googleVis

Pros: `r html_list(c("Interactive google maps"))`
Cons: `r html_list(c("No aesthetics control","No markdown","Riley doesn't work","Not dynamic"))`

```{r,echo=FALSE,message=FALSE,results='asis'}
google.location <- paste(dt$Lat, dt$Long, sep = ":")
dt.google <- data.frame(dt, google.location)
g.inter.map <- gvisMap(data = dt.google, locationvar = "google.location",tipvar = "Tip",
                       options=list(enableScrollWheel=T))
(g.inter.map$html$chart)
```

### leaflet from leafletR

Pros: `r html_list(c("Interactive"))`
Cons: `r html_list(c("Does not use Google maps","Not easy","No markdown","Runs on Riley, but does not show in window. Writes separate html file","Not dynamic"))`


```{r,echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
cat(paste(readLines(leaflet(data=toGeoJSON(dt,name="Andrew")))[23:53],collapse="\n"))

```

### plotGoogleMaps from plotGoogleMaps

Pros: `r html_list(c("Interactive","Only control color"))`
Cons: `r html_list(c("Not easy","No markdown","Runs on Riley, but does not show in window.","Not dynamic"))`


```{r,echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
df <- as.data.frame(dt)
coordinates(df) <- ~Long+Lat
proj4string(df) <- CRS('+proj=longlat')
#proj4string(df) <- CRS('+init=epsg:28992')

cat(plotGoogleMaps(df,zcol="Category",fillOpacity=0.9,colPalette=c("red","green","yellow"),openMap=F)[-1])

```

### ggmap from ggmap

Pros: `r html_list(c("Full aesthetics control","Markdown compatible","Riley2 works"))`
Cons: `r html_list(c("Interactive","Dynamic"))`


```{r,echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
#ggmap(get_map(location=c(median(dt$Long),median(dt$Lat)),zoom=12)) + geom_point(data=dt,aes(Long,Lat,col=Category,size=Speed_kt))

```


### D3 from scratch?

Pros: `r html_list(c("Interactive","Dynamic","Full aesthetics control","Markdown compatible","Riley2 works"))`
Cons: `r html_list(c("Building from scratch."))`


```{r,echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
#ggmap(get_map(location=c(median(dt$Long),median(dt$Lat)),zoom=12)) + geom_point(data=dt,aes(Long,Lat,col=Category,size=Speed_kt))

```






