---
title: "Untitled"
runtime: shiny
output: 
  html_document:
    self-contained: no
---

Try to insert tabs with formulas on each tab. Do formulas appear correctly after changing tabs?

```{r, echo=FALSE}
library(rmarkdown)
flowLayout(
  numericInput("rows","How many rows?",5),
  selectInput("letter", "Which letter?", LETTERS),
  sliderInput("value", "What value?", 0, 100, 50)
)
```

Some separation.

```{r, echo=FALSE}
fluidRow(
  column(width=5,numericInput("n_breaks","How many rows?",5),selectInput("letter2", "Which letter?", LETTERS)),
  column(width=5,offset=1,(sliderInput("value2", "What value?", 0, 100, 50)))
)
    observe({input$n_breaks;includeMarkdown(("txt.Rmd"))})
```


```{r,echo=FALSE,results='asis'}
renderText({
  text <- input$n_breaks
  cat("BLAH")
  includeMarkdown(render("txt.Rmd"))
})

```

Something between here.

```{r,echo=FALSE}


renderText({
  text <- "here is some text"
  if(input$n_breaks==5) {
    text <- paste("here is something $a^2+b^2=c^2$")  # HTML only compiles once, cannot be updated.
    write(text,"txt.Rmd")
  }
  else {
    text <- "file destroyed"
    write(text,"txt.Rmd")
  }
  text
})
```

space

```{r,echo=FALSE}
tabsetPanel(
  tabPanel("First",
    renderText({paste("This is a formula: $$a^2 + b^2 = c^2$$")}),
    renderPlot({plot(1)})
  ),
  tabPanel("Second",
    "This is some other text to print",
    renderText({paste("This is a formula: $$a^2 + b^2 = c^2$$")}),
    renderPlot({plot(1:10)})
  )
)
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package="shiny"),
  options=list(
    width="100%", height=550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



