---
title: "R Notebook"
output: html_notebook
---

```{r label, echo=FALSE, fig.height=10, fig.width=16, comment = FALSE, warning = FALSE}
# Integer Linear Programming Examples
library(data.table)
library(CMcCode)
library(htmlTable)
library(lpSolve)
library(readODS)
library(Rglpk)


Amat <- rbind(c(rep(1, 9), rep(-2, 9)),
              c(1, 1, 1, 0, 0, 0, 0, 0, 0, 1, 1, 1, 0, 0, 0, 0, 0, 0),
              c(0, 0, 0, 0, 0, 0, 1, 1, 1, 0, 0, 0, 0, 0, 0, 1, 1, 1),
              kronecker(rep(1, 6), c(1, 0, 0)),
              kronecker(rep(1, 6), c(0, 0, 1)),
              diag(18),
              diag(18)
)
colnames(Amat) <- paste0("x<sub>", 1:ncol(Amat), "</sub>")

bvec <- c(0, rep(11, 4), replace(rep(1, 18), c(5, 14), 0), replace(rep(3, 18), c(5, 14), 0))

cvec <- rep(1, 18)

ineq <- c("==", rep("==", 4), rep(">=", 18), rep("<=", 18))
ineq.p <- ifelse(ineq == "<=", "&le;", ifelse(ineq == ">=", "&ge;", "="))

ints <- 1:18

# Check dimensions
# dim(Amat); length(bvec); length(ineq); length(cvec)

#htmlTable(cbind(apply(matpaste(Amat, "x<sub>", col(Amat), "</sub>"), 1, paste, collapse = " + "), ineq, bvec), useViewer = T, css.cell = "padding-left: 1em; padding-right: 1em;", header = rep("", 3))
htmlTable(cbind(Amat, " " = ineq.p, Const = bvec), rnames = FALSE,
          css.cell = "padding-left: .5em; padding-right: .5em; font-size: 20px", 
          rgroup = c("Twice as many up as down", "11 per side", "At least 1 person per room", "No more than 3 people per room"), 
          n.rgroup = c(1, 4, 18, 18))

```



```{r, echo = FALSE}

minres <- lp(direction = "min", objective.in = cvec, const.mat = Amat, const.dir = ineq, const.rhs = bvec, int.vec = ints)
minres$solution

tab1 <- htmlTable(rbind(matrix(minres$solution[1:9], 3, 3), matrix(minres$solution[10:18], 3, 3)), rnames = F,
          css.cell = "padding: 30px; font-size: 25; border: solid;", n.rgroup = 3, rgroup = c("Top", "Bottom"),
          caption = "A Minimum Solution<br>Total People: 27")

maxres <- lp(direction = "max", objective.in = cvec, const.mat = Amat, const.dir = ineq, const.rhs = bvec, int.vec = ints)
maxres$solution

tab2 <- htmlTable(rbind(matrix(maxres$solution[1:9], 3, 3), matrix(maxres$solution[10:18], 3, 3)), rnames = F, 
          css.cell = "padding: 30px; font-size: 25; border: solid;", n.rgroup = 3, rgroup = c("Top", "Bottom"), 
          caption = "A Maximum Solution<br>Total People: 30")
```

<table class='gmisc_table' style='border-collapse: collapse; margin-top: 1em; margin-bottom: 1em; width:130%' >
<tr><td style = 'width:30%'>`r tab1`</td><td>&nbsp;</td><td>&nbsp;</td><td style = 'width:70%'>`r tab2`</td></tr></table>

Solution is not unique; however, multiple iterations give the same result.

```{r}
results <- NULL
for (i in 1:1e3) {
  minres <- lp(direction = "min", objective.in = cvec, const.mat = Amat, const.dir = ineq, const.rhs = bvec, int.vec = ints)
  results <- rbind(results, minres$solution)
}
apply(results, 2, function(x) length(unique(x)))
```


## Sudoku Solver

```{r, echo = FALSE}
n <- 3
initpos <- c(1, 2, 5, 6, 9, 13, 16, 18, 21, 23, 25, 30, 37, 38, 44, 45, 52, 57, 59, 61, 64, 66, 69, 73, 76, 77, 80, 81)
initval <- c(6, 8, 2, 4, 9, 1, 6, 2, 5, 9, 1, 2, 7, 3, 8, 1, 2, 7, 6, 4, 2, 9, 1, 3, 7, 8, 1, 6)
initmat <- do.call(rbind, lapply(initpos, 
                                 function(x) replace(rep(0, n^4), x, 1)))

tab <- matrix("", 9, 9)
tab[initpos] <- initval
cssmat <- matrix("border: solid; border-width: 1px; padding: 10px; font-size: 18px;", 9, 9)
cssmat <- matpaste(cssmat, ifelse(col(cssmat) %% 3 == 0, "border-right-width: 3px;", ""))
cssmat <- matpaste(cssmat, ifelse(col(cssmat) %% 3 == 1, "border-left-width: 3px;", ""))
cssmat <- matpaste(cssmat, ifelse(row(cssmat) %% 3 == 0, "border-bottom-width: 3px;", ""))
cssmat <- matpaste(cssmat, ifelse(row(cssmat) %% 3 == 1, "border-top-width: 3px;", ""))
htmlTable(t(tab), css.cell = cssmat)
```

First attempt: Optimize 81 integers between 1 and 9 to fit necessary criteria. Couldn't find constraints that ensure unique entries per column, row and block.

```{r, eval = FALSE, echo = FALSE}

# Only verifies sums, not unique entries ...

Amat <- rbind(rep(1, n^4),
              K(diag(n^2), t(rep(1, n^2))),                           # Row sums all equal
              K(t(rep(1, n^2)), diag(n^2)),                           # Column sums all equal
              K(diag(n), K(t(rep(1, n)), K(diag(n), t(rep(1, 3))))),  # Block sums all equal
              diag(n^4),                                              # Each entry is at least 1
              diag(n^4),                                              # Each entry is at most 9
                                                                      # Ensure each entry is unique ?!?!
              initmat )                                               # Given entries are fixed
colnames(Amat) <- paste0("x<sub>", 1:ncol(Amat), "</sub>")

bvec <- c(n^2*sum(1:n^2), rep(sum(1:n^2), 3*n^2), rep(1, n^4), rep(n^2, n^4), initval)

cvec <- rep(1, n^4)

ineq <- c("=", rep("=", 3*n^2), rep(">", n^4), rep("<", n^4), rep("=", length(initval)))
ineq.p <- ifelse(ineq == "<", "&le;", ifelse(ineq == ">", "&ge;", "="))

ints <- 1:n^4

res <- lp(direction = "min", objective.in = cvec, const.mat = Amat, const.dir = ineq, const.rhs = bvec, int.vec = ints)
res$solution
# sums are correct, but uniqueness is not obtained. 

```

## Sudoku Solver 2

Second attempt: Optimize 729 binary indicators, one for each combination of position and entry.

```{r, echo = FALSE}

# initpos <- c(1, 2, 5, 6, 9, 13, 16, 18, 21, 23, 25, 30, 37, 38, 44, 45, 52, 57, 59, 61, 64, 66, 69, 73, 76, 77, 80, 81)
# initval <- c(6, 8, 2, 4, 9, 1, 6, 2, 5, 9, 1, 2, 7, 3, 8, 1, 2, 7, 6, 4, 2, 9, 1, 3, 7, 8, 1, 6)

solve <- function(mat = NULL, initpos = NULL, initval = NULL, n = 3, byrow = T, print = T) {
  
  if(is.null(mat)) {
    mat <- matrix("", n^2, n^2)
    mat[initpos] <- initval
    if(byrow)
      mat <- t(mat)
  }
  
  mat <- as.matrix(mat)
  initpos = which(!mat %in% c("", 0, NA))
  initval = as.numeric(mat[initpos])
  initmat <- do.call(rbind, lapply(n^4*(initval - 1) + initpos, 
                                   function(x) replace(rep(0, n^6), x, 1)))
  
  K <- kronecker
  Amat <- rbind(K(diag(n^4), t(rep(1, n^2))),                                         # One of each number per column
                K(diag(n^2), K(t(rep(1, n^2)), diag(n^2))),                           # One of each number per row
                K(diag(n^2), K(K(diag(n), t(rep(1, n))), K(diag(n), t(rep(1, n))))),  # One of each number per block
                K(t(rep(1, n^2)), diag(n^4)),                                         # One number per position
                initmat )
  
  bvec <- rep(1, nrow(Amat))
  ineq <- rep("==", nrow(Amat))
  cvec <- rep(1, ncol(Amat))
  ints <- 1:n^6
  
  # Check that dimensions match
  # dim(Amat); length(bvec); length(ineq); length(cvec)
  
  # t0 <- proc.time()
  # res <- lp(direction = "min", objective.in = cvec, const.mat = Amat, const.dir = ineq, const.rhs = bvec, binary.vec = ints)
  # t1 <- proc.time()   # Did not scomplete in 30 minutes.
  
  res_glpk <- Rglpk_solve_LP(obj = cvec, mat = Amat, dir = ineq, rhs = bvec, types = "B", max = F)
  names(res_glpk)
  res_glpk$solution
  
  arr <- array(res_glpk$solution, dim = rep(n^2, 3))
  solution <- matrix(apply(arr, 1:2, function(x) which(x == 1)), n^2, n^2)
  
  cssmat <- matrix("border: solid; border-width: 1px; border-color: black; color: blue; padding: 10px; font-size: 18px;", 9, 9)
  cssmat <- matpaste(cssmat, ifelse(col(cssmat) %% 3 == 0, "border-right-width: 3px;", ""))
  cssmat <- matpaste(cssmat, ifelse(col(cssmat) %% 3 == 1, "border-left-width: 3px;", ""))
  cssmat <- matpaste(cssmat, ifelse(row(cssmat) %% 3 == 0, "border-bottom-width: 3px;", ""))
  cssmat <- matpaste(cssmat, ifelse(row(cssmat) %% 3 == 1, "border-top-width: 3px;", ""))
  cssmat[initpos] <- paste(cssmat[initpos], "color: black; font-weight: 900")
  solution_tab <- htmlTable(solution, css.cell = cssmat)
  
  row.names(mat) <- colnames(mat) <- NULL
  initial_tab <- htmlTable(mat, css.cell = cssmat)
  
  list(solution = solution, solution_tab = solution_tab, initial = mat, initial_tab = initial_tab)
}

puzzles <- read.ods("puzzle.ods")

puzzle1 <- solve(puzzles[[1]])
puzzle2 <- solve(puzzles[[2]])
puzzle3 <- solve(puzzles[[3]])

```

<table class='gmisc_table' style='border-collapse: collapse; margin-top: 1em; margin-bottom: 1em; width:130%' >
<tr><td style = 'width:30%'>`r puzzle1$initial_tab`</td><td>&nbsp;</td><td>&nbsp;</td><td style = 'width:70%'>`r puzzle1$solution_tab`</td></tr></table>

```{r, options}
# Read puzzle from .ods file.

puzzles <- read.ods("puzzle.ods")


```
