---
title: "Propensity Score Matching"
output: html_document
---

```{r setup, include=FALSE}
library(CMcCode)
library(data.table)
library(ggplot2)
library(grid)
library(gridExtra)
library(gtools)
library(htmlTable)


environment(reglines) <- globalenv()

padding <- function(n = 1) paste0("padding-right: ", n, "em; padding-left: ", n, "em;")
```


## Linear Regression

### Continuous explanatory variable 

```{r, echo = F, warning = F}

n <- 1e3
dt <- data.table(x = rnorm(n, 69))[, y := rnorm(.N, 26.5 + 0.6*x, 1.8)]

ggplot(dt, aes(x, y)) + geom_point()

```

```{r, echo = F, warning = F}

mod <- lm(y~x, data = dt)
reglines(mod)

html_summary(mod)

```

$$ Height = `r round(coef(mod), 1)[1]` + `r round(coef(mod), 1)[2]`ParentHeight$$

```{r, echo = F, warning = F}
n <- 40
dt <- data.table(x = rnorm(n))[, y := rnorm(.N, x)]
dt$ID <- factor(1:nrow(dt))
mod <- lm(y~x, data = dt)
dt$Pred <- predict(mod)

dt2 <- melt(dt, id.vars = c("x", "ID"), value.name = "y", variable.name = "type")

ggplot(dt, aes(x, y)) + geom_point() + geom_line(data = dt2, aes(x, y, group = ID)) + geom_abline(aes(slope = coef(mod)[2], intercept = coef(mod)[1]))
```

### Categorical explantory ariable

```{r, echo = F, warning = F}
n <- 500
dt <- data.table(x = rbinom(n, 1, .5))[, Height := rnorm(.N, 5*x + 64, 2.7)][, Gender := ifelse(x == 1, "Male", "Female")]

ggplot(dt, aes(Gender, Height)) + geom_point() + labs(y = "Height (in)")

```


```{r, echo = F, warning = F}

mod <- lm(Height~x, data = dt)
coefs <- coef(mod)

ggplot(dt, aes(x, Height)) + geom_point() + geom_abline(aes(slope = coefs[2], intercept = coefs[1]))
html_summary(mod)
```

$$ Height = `r round(coef(mod), 1)[1]` + `r round(coef(mod), 1)[2]`Male$$

### Multiple explanatory variables

```{r, echo = F, warning = F}

n <- 500
dt <- data.table(x = rnorm(n, 65), g = rbinom(n, 1, .5))[, Height := rnorm(.N, 26.5 + 0.6*x + 4*g, 1.6)][, Gender := ifelse(g==1, "Male", "Female")]

mod <- lm(Height ~ x, data = dt)
ggplot(dt, aes(x, Height)) + geom_point() + geom_abline(aes(slope = coef(mod)[2], intercept = coef(mod)[1]))


```

```{r, echo = F, warning = F}

mod <- lm(Height ~ x + g, data = dt)
ggplot(dt, aes(x, Height, col = Gender)) + geom_point() + 
  geom_abline(data = data.frame(slp = coef(mod)[2], int = cumsum(coef(mod)[-2]), Gender = c("Female", "Male")), 
              aes(slope = slp, intercept = int, col = Gender))

```

$$ Height = `r round(coef(mod), 1)[1]` + `r round(coef(mod), 1)[2]`ParentHeight + `r round(coef(mod), 1)[3]`Male$$

... and they can easily get more complicated.

```{r, echo = F}
coef <- c(2.1, -0.8, 1.2, -0.9, 2, -3)
dt <- data.table(x1 = rnorm(n))
dt[, x2 := rnorm(.N, -.8 + .7*x1)]
dt[, Group := letters[rbinom(.N, 2, .5) + 1]]
dt[, x3 := rnorm(.N, 1.2 + 1.1*x1 + .5*x2)]
dt[, Mean := cbind(x1, x2, x3, x1*x2, factor(Group), as.numeric(factor(Group))*x3) %*% coef]
dt[, y := rnorm(.N, Mean, .5)]
mod <- lm(y ~ (x1 + x2 + x3)*Group + x1*x2, data = dt)
reglines(mod)
```

## Experimental Design {.tabset .tabset-pills}

### What is the most important part of any experiemental design?

### Answer

#### Randomization!

Without proper randomization, all results of an experiment may be meaningless. 

Randomization is important at two points in every experiment
`r html_list(c("Sampling", "Assignment"))`

## Example

A researcher is trying to measure the effect of AP courses taken during the last year of high school on SAT scores. In an analysis of 1,000 students, a linear regression is set up with SAT scores as a response variable and whether or not the student took at least one AP course as an explantory variable (AP = 1 if the student took an AP course, AP = 0 otherwise). The proposed model is:

$$ SAT = \beta_0 + \beta_1 AP$$

### Analysis {.tabset .tabset-pills}

#### Method 1

```{r, echo = FALSE}

n <- 1e3
dt <- data.table(GPA = rbeta(n, 3, 1.2)*3 + 1)[, AP := ifelse(GPA > 3.6, 1, ifelse(GPA < 2.4, 0, rbinom(.N, 1, (GPA - 2.4)/1.2)))]
dt$GPA <- round(dt$GPA, 2)
dt$SAT <- rnorm(n, ifelse(dt$GPA > 3.6, 600, ifelse(dt$GPA < 2.4, 600, 300))*dt$GPA + 100*dt$AP + 
                  ifelse(dt$GPA > 3.6, -500, ifelse(dt$GPA < 2.4, -200, 500)), 100)
dt$SAT <- round(dt$SAT)
# dt[, list(mean(SAT), sd(SAT))]
dt$APfact <- factor(dt$AP, levels = 0:1, labels = c("No AP Course", "AP Course"))

ggplot(dt, aes(APfact, SAT)) + geom_boxplot()

mod <- lm(SAT ~ AP, dat = dt)
#mod <- lm(SAT ~ AP + GPA, dat = dt[GPA > 2.4 & GPA < 3.6])
html_summary(mod)
```

$$ SAT = `r round(coef(mod), 1)[1]` + `r round(coef(mod), 1)[2]`AP$$

Problems?

#### Method 2

Realizing there was no randomization in determining which students would take an AP class, the researcher decides to rerun the analysis but considers junior year GPA as an explanatory variable, in addition to the AP course indicator.

```{r, echo = FALSE}

ggplot(dt, aes(GPA, SAT, col = APfact)) + geom_point() + labs(col = "")

mod <- lm(SAT ~ AP + GPA, dat = dt)
html_summary(mod)
```

$$ SAT = `r round(coef(mod), 1)[1]` + `r round(coef(mod), 1)[2]`AP + `r round(coef(mod), 1)[3]`GPA$$

Problem solved ... ?

### Other things to consider

#### Different ranges of GPA values

```{r, echo = FALSE, warning = FALSE, message = FALSE}


ggplot(dt, aes(GPA, fill = factor(AP))) + geom_histogram(position = "dodge")
ggplot(dt, aes(GPA, fill = factor(AP))) + geom_density(alpha = .3)

ggplot(dt, aes(GPA, SAT, fill = APfact, col = GPA > 2.4 & GPA < 3.6)) + geom_point(pch = 21) + labs(fill = "") + scale_color_manual(values = c("white", "black"))

ggplot(dt[GPA > 2.4 & GPA < 3.6], aes(GPA, SAT, col = APfact)) + geom_point() + labs(col = "")

mod <- lm(SAT ~ AP + GPA, dat = dt[GPA < 3.6 & GPA > 2.4])
html_summary(mod)
```

#### GPA influenced both SAT scores and whether AP courses were taken

```{r, echo = FALSE}

mod <- lm(AP ~ GPA, data = dt)
html_summary(mod)

dt$Pred <- predict(mod)

htmlTable(head(dt[, list(GPA, AP = APfact, SAT, Pred = round(Pred, 3))], 15), rnames = F, css.cell = padding(1), align = "rlrr")
```


## Matching

The goal of matching is to simulate randomization by selecting individuals from both groups that are similar with respect to a certain set of characteristics. By matching on certain variables, we hopefully remove any effect of those variables on the group membership (this also changes the effect of those variables on the response variable)

Matching Procedure

* Find a distance metric
* Choose matching method
* Evaluate matches
* Perform analysis

### Distance metric

* Exact
* Mahalanobis
* Propensity score







